Sharing statistics for SPARQL Federation optimization, with emphasis on benchmark quality

### 2012-01-12 eswc-2012/phdsymp.tex
well-designed patterns
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@76 43ef1190-90fd-4dd7-a769-d0121f02b80e

A recent scientific treatment of the current specification, is in
\cite{springerlink:10.1007/978-3-642-21064-8-1}. In this paper, the
authors also show an optimization strategy based on execution order of
different algebraic elements. % TODO: wording here

@@

This part of the thesis seeks to use advances from statistical
experimental planning to improve accuracy and dependability of
benchmarking. Already in 1926, Ronald~Fischer noted that complex
experiments can be much more efficient than simple
ones\footnote{Cited in
  http://en.wikipedia.org/wiki/Factorial\_experiment}, starting the
experimental design field. One of the simpler designs is ``fractional
factorial design'', in which several ``factors'' are studied by using
different ``treatments''. In terms of SPARQL execution, one factor can
for example be the level of nested \texttt{OPTIONAL}s, number of
triples in a basic graph pattern, etc, and the treatment is to vary
these numbers. The key to understanding why this can be efficient is
that these variations need not occur in the same experiment. Thus,
many combinations of factors can be studied by carefully designing
queries to cover different factors, and a formalism called
``resolution'' has been developed to classify how well this has been
achieved.

Fractional factorial design is covered in elementary text-books in
statistics, and I intend to go further into the statistical literature
to see if there is methodology that is even better suited to the
problem. Armed with this, I hope to use amongst other things, the
complexity analysis of SPARQL (see
e.g. \cite{Schmidt:2010:FSQ:1804669.1804675}) to find suitable factors
to see if my admittedly bold proposition that it is possible to design
a benchmark to cover all realistic cases.

>>>>

A recent scientific treatment of the current specification, is in
\cite{springerlink:10.1007/978-3-642-21064-8-1}. In this paper, the
authors also show an optimization strategy based on execution order of
so-called well-designed patterns.

@@

This part of the thesis seeks to use advances from statistical
experimental planning to improve accuracy and dependability of
benchmarking. Already in 1926, Ronald~Fischer noted that complex
experiments can be much more efficient than simple ones\footnote{Cited
  in http://en.wikipedia.org/wiki/Factorial\_experiment}, starting the
experimental design field. One of the simpler designs is ``fractional
factorial design'', in which several ``factors'' are studied by using
different ``treatments''. In terms of SPARQL execution, the SPARQL
engine is clearly a factor, but also for example the nestedness of
\texttt{OPTIONAL}s can be a factor, number of triples in a basic graph
pattern, etc, and the treatment is to vary these numbers to different
``levels''. The key to understanding why this can be efficient is that
these variations need not occur in the same experiment. Thus, many
combinations of factors can be studied by carefully designing queries
to cover different factors, and a formalism called ``resolution'' has
been developed to classify how well this has been achieved.

Fractional factorial design is covered in elementary text-books in
statistics but is most likely inadequate for this purpose, so I
intend to go further into the statistical literature to see if there
is methodology that is even better suited to the problem. Armed with
this, I hope to use amongst other things, the complexity analysis of
SPARQL (see e.g. \cite{Schmidt:2010:FSQ:1804669.1804675}) to find
suitable factors to see if my admittedly bold proposition that it is
possible to design a benchmark to cover all realistic cases.
